---
title: "CodesRUs_midterm"
authors: "Daniel Diaz, Ganesh Ghimire, Rose Mesina, and Michael Womble"
date: "March 9, 2022"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
library(ggplot2)
library(ggsoccer)
library(dplyr)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

# R Markdown - Exploratory Data Analysis of Soccer Event Data Codes

**SMART Question:** Based on the matches played during the 2017/2018 season for Europe’s top five leagues, the 2016 European Championship, and the 2018 World Cup, which match variables (ex: shot location, assist location, assist duration, etc…) result in the highest probability of a shot on goal being successful, result in a goal?

## Importing Data:

**Dataset:** The dataset contains all matches played in Europe’s top five leagues during the 2017/2018 season (380 matches per league for a total of 1900 matches, producing 643,149 match events), the 2016 European Championship (51 matches, producing 78,139 match events), and the 2018 World Cup (64 matches, 101,758 match events). It was obtained from the work posted by Luca Pappalardo on Figshare.

The data is imported by league and then compiled into a single event_data dataframe, giving a total of 3,251,294 events and 19 variables
```{r load data, results='markup', out.width='50%'}
#Load Event data from each competition as a dataframe
england<-data.frame(read.csv('./datasets/events_England_new.csv'))
england$competition <- "England"

italy<-data.frame(read.csv('./datasets/events_Italy_new.csv'))
italy$competition <- "Italy"

germany<-data.frame(read.csv('./datasets/events_Germany_new.csv'))
germany$competition <- "Germany"

france<-data.frame(read.csv('./datasets/events_France_new.csv'))
france$competition <- "France"

spain<-data.frame(read.csv('./datasets/events_Spain_new.csv'))
spain$competition <- "Spain"

european<-data.frame(read.csv('./datasets/events_European_Championship_new.csv'))
european$competition <- "European"

worldcup<-data.frame(read.csv('./datasets/events_World_Cup_new.csv'))
worldcup$competition <- "World Cup"
                     
#Append dataframes into a single dataframe from which to draw.
event_data<-rbind(england, italy, germany, france, spain, european, worldcup)
#str(event_data)
xkabledplyhead(event_data)
```
**Table 1.** First five rows of event_data data frame.

## Cleaning Data:
X and Y coordinates had been imported as characters. We therefore converted them to numeric to allow us to work with them. Event Name and Subevent Name are have also been converted to factor. 

```{r convert variables}
#convert x and y coordinate columns from characters to integer values
event_data$x.End <- as.numeric(event_data$x.End)
event_data$y.End <- as.numeric(event_data$y.End)

event_data$x.Start <- as.numeric(event_data$x.Start)
event_data$y.Start <- as.numeric(event_data$y.Start)

#convert event, subevent, tags, players, team labels, etc... to factors
event_data$Event.Name=as.factor(event_data$Event.Name)
event_data$Subevent.Name=as.factor(event_data$Subevent.Name)
event_data$Match.Period=as.factor(event_data$Match.Period)
event_data$Match.Date=as.factor(event_data$Match.Date)
event_data$Player=as.factor(event_data$Player)
event_data$Team.Name=as.factor(event_data$Team.Name)
event_data$Tag.1=as.factor(event_data$Tag.1)
event_data$Tag.2=as.factor(event_data$Tag.2)
event_data$Tag.3=as.factor(event_data$Tag.3)
event_data$Tag.4=as.factor(event_data$Tag.4)
event_data$Tag.5=as.factor(event_data$Tag.5)
event_data$Tag.6=as.factor(event_data$Tag.6)

str(event_data)
```

## Calculating and Adding Event Distance Variable:
One variable we are interested in looking at is the distance of assists and key passes and whether this plays a role in the success of the pass and whether it results in a successful shot on goal or not. We have to calculate distance for each event using the x,y-coordinates given in the data set.

```{r calculate distance}

event_data<-event_data %>% mutate(dist = sqrt((x.End-x.Start)^2+(y.End-y.Start)^2))

event_data$dist <- as.numeric(event_data$dist)

summary(event_data$dist)
str(event_data)
```

## Subsetting Data:
Separate data frames are created for each of the following events of interest: 

1. Shots: unsuccessful shots on goal  
2. goals: successful shots on goal  
3. assists: pass leading to a goal  
4. key passes: pass leading to unsuccessful scoring opportunity  

```{r event datasets}
#Collect shots, goals, assists, and key passes into their own dataframes
shots<-subset(event_data, Event.Name=='Shot')
shots<-subset(shots, Tag.1!='Goal')
goals<-subset(event_data, Tag.1=='Goal')
goals<-subset(goals, Event.Name!='Save attempt')
assists<-subset(event_data, Tag.1=='Assist')
key_passes<-subset(event_data, Tag.1=='Key pass')

head(shots)
head(goals)
head(assists)
head(key_passes)

```

## Checking Each Variable for Outliers:
We checked each variable of interest for outliers and removed them. **Figure 1.**, below, shows the outlier check for the assist distance variable. Many outliers are seen at the high end of the histogram resulting in a skewed distribution. This is also seen in both the q-q plot and the boxplot. After removing the outliers, a better normal distribution is seen. This was done to the following variables of interest:  

1. Assists distance
2. Assists x & y-coordinates
3. Key passes distance
4. Key passes x & y-coordinates
5. Shots x & y-coordinates
6. Goals x & y-coordinates

```{r checking for outliers}
#Set to TRUE of all plots for each varialbe want to be seen.
show_all_plots=FALSE

assists_clean<-outlierKD2(assists, dist, rm=TRUE, boxplt = TRUE, histogram = TRUE, qqplt = TRUE)
assists_clean<-na.omit(assists_clean)

assists_clean_x<-outlierKD2(assists, x.Start, rm=TRUE, 
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
assists_clean_x<-na.omit(assists_clean_x)

assists_clean_y<-outlierKD2(assists, y.Start, rm=TRUE, 
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
assists_clean_y<-na.omit(assists_clean_y)

key_passes_clean<-outlierKD2(key_passes, dist, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
key_passes_clean<-na.omit(key_passes_clean)

key_passes_clean_x<-outlierKD2(key_passes, x.Start, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
key_passes_clean_x<-na.omit(key_passes_clean_x)

key_passes_clean_y<-outlierKD2(key_passes, y.Start, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
key_passes_clean_y<-na.omit(key_passes_clean_y)

shots_clean<-outlierKD2(shots, dist, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
shots_clean<-na.omit(shots_clean)

shots_clean_x<-outlierKD2(shots, x.Start, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
shots_clean_x<-na.omit(shots_clean_x)

shots_clean_y<-outlierKD2(shots, y.Start, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
shots_clean_y<-na.omit(shots_clean_y)

goals_clean<-outlierKD2(goals, dist, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
goals_clean<-na.omit(goals_clean)

goals_clean_x<-outlierKD2(goals, x.Start, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
goals_clean_x<-na.omit(goals_clean_x)

goals_clean_y<-outlierKD2(goals, y.Start, rm=TRUE,  
                            boxplt = show_all_plots, 
                            histogram = show_all_plots, 
                            qqplt = show_all_plots)
goals_clean_y<-na.omit(goals_clean_y)

```
**Figure 1.** Outlier check for assist distance variable.  

## Visualizing Variable Distributions.

Normalized histograms of previously specified variables are plotted and fitted with normal distributions. Upon reviewing the plots, it appears that there are some differences between some variables. We will continue by performing some hypothesis tests. Some of the distributions are also not normal.
```{r viz dist, fig.show='hold', out.width='33%'}

#Histograms of Assist and Key Pass Distance

ggplot()+
  geom_histogram(data=assists_clean, aes(x=dist, y=..density..), color='Black', fill='Green', alpha=0.3)+
  
  geom_histogram(data=key_passes_clean, aes(x=dist, y=..density..), color='Black', fill='Red', alpha=0.3)+
  
  stat_function(fun = dnorm, args = list(mean = mean(assists_clean$dist), sd = sd(assists_clean$dist)), color='dark green', fill='green', size=1)+
  
  stat_function(fun = dnorm, args = list(mean = mean(key_passes_clean$dist), sd = sd(key_passes_clean$dist)), color='dark red', fill='green', size=1)+
  
  labs(title="Distance of Assists and Key Passes with Normal Distribution Fit", x="Distance", y="Normalized Count")+
  
  scale_color_manual(name='Key',values=c('Green', 'Red'), labels=c('Assists','Key Passes'))+
  theme(panel.background = element_blank(),
        panel.border=element_rect(color='black', fill=NA),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14))

ggsave("assist_keypass_dist.jpeg")

#Histograms of Assist and Key Pass x Coordinates

ggplot()+
  geom_histogram(data=(assists_clean_x), aes(x=x.Start, y=..density..), color='Black', fill='Green', alpha=0.3)+
  
  geom_histogram(data=(key_passes_clean_x), aes(x=x.Start, y=..density..), color='Black', fill='Red', alpha=0.3)+
  
  stat_function(fun = dnorm, args = list(mean = mean(assists_clean_x$x.Start), sd = sd(assists_clean_x$x.Start)), color='dark green', fill='green', size=1)+
  
  stat_function(fun = dnorm, args = list(mean = mean(key_passes_clean_x$x.Start), sd = sd(key_passes_clean_x$x.Start)), color='dark red', fill='red', size=1)+
  
  labs(title="X-coordinates of Key Passes & Assists with Normal Distribution Fits", x="x-coordinate", y="Normalized Count")+
  
  scale_color_manual(name='Key',values=c('Green', 'Red'), labels=c('Assists','Key Passes'))+
  theme(panel.background = element_blank(),
        panel.border=element_rect(color='black', fill=NA),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14))

ggsave("assist_keypass_x.jpeg")

#Histograms of Assist and Key Pass y Coordinates

ggplot()+
  geom_histogram(data=(assists_clean_y), aes(x=y.Start, y=..density..), color='Black', fill='Green', alpha=0.3)+
  
  geom_histogram(data=(key_passes_clean_y), aes(x=y.Start, y=..density..), color='Black', fill='Red', alpha=0.3)+
  
  stat_function(fun = dnorm, args = list(mean = mean(assists_clean_y$y.Start), sd = sd(assists_clean_y$y.Start)), color='dark green', fill='green', size=1)+
  
  stat_function(fun = dnorm, args = list(mean = mean(key_passes_clean_y$y.Start), sd = sd(key_passes_clean_y$y.Start)), color='dark red', fill='red', size=1)+
  
  labs(title="Y-coordinates of Key Passes & Assists with Normal Distribution Fits", x="y-coordinate", y="Normalized Count")+
  
  scale_color_manual(name='Key',values=c('Green', 'Red'), labels=c('Assists','Key Passes'))+
  theme(panel.background = element_blank(),
        panel.border=element_rect(color='black', fill=NA),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14))

ggsave("assist_keypass_y.jpeg")
```
**Figure 2.** Histograms and normal distribution fits of assists (in green) and key passes (in red) (Left) pass distance, (Middle) x-coordinate, (Right) y-coordinate.  
```{r histogram of shots and goals, fig.show='hold', out.width='33%'}
#Histograms of shots and goals x coordinates
ggplot()+
  geom_histogram(data=(shots_clean), aes(x=dist, y=..density..), color='Black', fill='Red', alpha=0.3)+
  
  geom_histogram(data=(goals_clean), aes(x=dist, y=..density..), color='Black', fill='Blue', alpha=0.3)+
  
  stat_function(fun = dnorm, args = list(mean = mean(shots_clean$dist), sd = sd(shots_clean$dist)), color='dark red', fill='red', size=1)+
  
  stat_function(fun = dnorm, args = list(mean = mean(goals_clean$dist), sd = sd(goals_clean$dist)), color='dark blue', fill='blue', size=1)+
  
  labs(title="Distance of Shots & Goals with Normal Distribution Fits", x="Shot Distance", y="Normalized Count")+
  
  scale_color_manual(name='Key',values=c('Blue', 'Red'), labels=c('Goals','Shots'))+
  theme(panel.background = element_blank(),
        panel.border=element_rect(color='black', fill=NA),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14))

ggsave("shots_goals_dist.jpeg")


ggplot()+
  geom_histogram(data=(shots_clean_x), aes(x=x.Start, y=..density..), color='Black', fill='Red', alpha=0.3)+
  
  geom_histogram(data=(goals_clean_x), aes(x=x.Start, y=..density..), color='Black', fill='Blue', alpha=0.3)+
  
  stat_function(fun = dnorm, args = list(mean = mean(shots_clean_x$x.Start), sd = sd(shots_clean_x$x.Start)), color='dark red', fill='red', size=1)+
  
  stat_function(fun = dnorm, args = list(mean = mean(goals_clean_x$x.Start), sd = sd(goals_clean_x$x.Start)), color='dark blue', fill='blue', size=1)+
  
  labs(title="X-coordinates of Shots & Goals with Normal Distribution Fits", x="x-coordinate", y="Normalized Count")+
  
  scale_color_manual(name='Key',values=c('Blue', 'Red'), labels=c('Goals','Shots'))+
  theme(panel.background = element_blank(),
        panel.border=element_rect(color='black', fill=NA),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14))

ggsave("shots_goals_x.jpeg")

#Histograms of shots and goals y coordinates

ggplot()+
  geom_histogram(data=shots_clean_y, aes(x=y.Start, y=..density..), color='Black', fill='Red', alpha=0.3)+
  
  geom_histogram(data=goals_clean_y, aes(x=y.Start, y=..density..), color='Black', fill='Blue', alpha=0.3)+
  
  stat_function(fun = dnorm, args = list(mean = mean(shots_clean_y$y.Start), sd = sd(shots_clean_y$y.Start)), color='dark red', fill='red', size=1)+
  
  stat_function(fun = dnorm, args = list(mean = mean(goals_clean_y$y.Start), sd = sd(goals_clean_y$y.Start)), color='dark blue', fill='blue', size=1)+
  
  labs(title="Y-coordinates of Shots & Goals with Normal Distribution Fits", x="y-coordinate", y="Normalized Count")+
  
  scale_color_manual(name='Key',values=c('Blue', 'Red'), labels=c('Goals','Shots'))+
  theme(panel.background = element_blank(),
        panel.border=element_rect(color='black', fill=NA),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14))

ggsave("shots_goals_y.jpeg")
```
**Figure 3.** Histograms and normal distribution fits of goals (in blue) and shots (in red) (Left) pass distance, (Middle) x-coordinate, (Right) y-coordinate.

## T-tests

Based on the t-tests, the following variables have significantly different means at the 95% level.

1. Assist v. Key Pass Distance (higher mean for key pass)
2. Goal v. Shot Distance (higher mean for goals)
3. Goal v. Shot x.Start (higher mean for goals)
4. Goal v. Shot y.Start (higher mean for goals)
```{r hypothesis testing}

#Distance

ttest_assist_v_keypass_dist <- t.test(assists_clean$dist, key_passes_clean$dist)

capture.output(ttest_assist_v_keypass_dist, file = "ttests.txt", append = T)

ttest_goals_v_shots_dist <- t.test(goals_clean$dist, shots_clean$dist)

capture.output(ttest_goals_v_shots_dist, file = "ttests.txt", append = T)

#Starting X-coordinate

ttest_assist_v_keypass_x <- t.test(assists_clean_x$x.Start, key_passes_clean_x$x.Start)

capture.output(ttest_assist_v_keypass_x, file = "ttests.txt", append = T)

ttest_goals_v_shots_x <- t.test(goals_clean_x$x.Start, shots_clean$x.Start)

capture.output(ttest_goals_v_shots_x, file = "ttests.txt", append = T)

#Starting Y-coordinate

ttest_assist_v_keypass_y <- t.test(assists_clean_y$y.Start, key_passes_clean_y$y.Start)

capture.output(ttest_assist_v_keypass_y, file = "ttests.txt", append = T)

ttest_goals_v_shots_y <- t.test(goals_clean_y$y.Start, shots_clean$y.Start)

capture.output(ttest_goals_v_shots_y, file = "ttests.txt", append = T)

```

## Success Probabilities vs Pitch Position:

Success probabilities of shots and assists/key passes are calculated as functions of event distance, and x & y-coordinates. These plots show that all these variables play a role in determining the probability of success of a shot resulting in a goal.
```{r Success Probabilites, fig.show='hold', out.width='33%'}
#Assist:Key pass ratio
assist_density_length<-density(assists_clean$dist,
                               n=100, from = 0, to = 100)
key_pass_density_length<-density(key_passes_clean$dist,
                                 n=100, from = 0, to = 100)

assist_density_length$y=assist_density_length$n/sum(assist_density_length$y)*assist_density_length$y
key_pass_density_length$y=key_pass_density_length$n/sum(key_pass_density_length$y)*key_pass_density_length$y
total_keyassists=assist_density_length$y+key_pass_density_length$y
assist_prob_length=assist_density_length$y/total_keyassists

plot(assist_prob_length, ylim=c(0,1),
     xlab='Pass Distance',
     ylab='Goal Probability',
     main = 'Probability of pass resulting in a goal vs distance of pass')

#Assist:key pass ratio x-coordinate
assist_density_x<-density(assists_clean_x$x.Start,
                          n=100, from = 0, to = 100)
key_pass_density_x<-density(key_passes_clean_x$x.Start,
                            n=100, from = 0, to = 100)

assist_density_x$y=assist_density_x$n/sum(assist_density_x$y)*assist_density_x$y
key_pass_density_x$y=key_pass_density_x$n/sum(key_pass_density_x$y)*key_pass_density_x$y
total_keyassists_x=assist_density_x$y+key_pass_density_x$y
assist_prob_x=assist_density_x$y/total_keyassists_x

plot(assist_prob_x, 
     xlim=c(50, 100), ylim=c(0,1),
     xlab='Field Length Percentage',
     ylab='Goal Probability',
     main = 'Probability of pass resulting in a goal vs x-coordinate of pass')

#Assist:key pass ratio y-coordinate
assist_density_y<-density(assists_clean_y$y.Start,
                          n=100, from = 0, to = 100)
key_pass_density_y<-density(key_passes_clean_y$y.Start,
                            n=100, from = 0, to = 100)

assist_density_y$y=assist_density_y$n/sum(assist_density_y$y)*assist_density_y$y
key_pass_density_y$y=key_pass_density_y$n/sum(key_pass_density_y$y)*key_pass_density_y$y
total_keyassists_y=assist_density_y$y+key_pass_density_y$y
assist_prob_y=assist_density_y$y/total_keyassists_y

plot(assist_prob_y, ylim=c(0,1),
     xlab='Field Width Percentage',
     ylab='Goal Probability',
     main = 'Probability of pass resulting in a goal vs y-coordinate of pass')

#Goal percentage shot distance
goal_density_dist<-density(goals_clean$dist,
                        n=100, from=0, to=100)

shots_density_dist<-density(shots_clean$dist,
                         n=100, from=0, to=100)

goal_density_dist$y=goal_density_dist$n/sum(goal_density_dist$y)*goal_density_dist$y
shots_density_dist$y<-shots_density_dist$n/sum(shots_density_dist$y)*shots_density_dist$y
total_shots_dist<-goal_density_dist$y+shots_density_dist$y
goal_shot_prob_dist<-(goal_density_dist$y/total_shots_dist)

plot(goal_shot_prob_dist, 
     xlim=c(0, 100), ylim=c(0,1),
     xlab='Shot Distance',
     ylab='Goal Probability',
     main = 'Probability of Shot resulting in a goal vs distance of shot')

#Goal percentage x-coordinate
goal_density_x<-density(goals_clean_x$x.Start,
                        n=100, from=0, to=100)

shots_density_x<-density(shots_clean_x$x.Start,
                         n=100, from=0, to=100)

goal_density_x$y=goal_density_x$n/sum(goal_density_x$y)*goal_density_x$y
shots_density_x$y<-shots_density_x$n/sum(shots_density_x$y)*shots_density_x$y
total_shots_x<-goal_density_x$y+shots_density_x$y
goal_shot_prob_x<-(goal_density_x$y/total_shots_x)

plot(goal_shot_prob_x, 
     xlim=c(60, 100), ylim=c(0,1),
     xlab='Field Length Percentage',
     ylab='Goal Probability',
     main = 'Probability of Shot resulting in a goal vs x-coordinate of shot')

#Goal percentage y-coordinate
goal_density_y<-density(goals_clean$y.Start,
                        n=100, from=5, to=95)
shots_density_y<-density(shots_clean_y$y.Start,
                         n=100, from=5, to=95)

goal_density_y$y=goal_density_y$n/sum(goal_density_y$y)*goal_density_y$y
shots_density_y$y=shots_density_y$n/sum(shots_density_y$y)*shots_density_y$y
total_shots_y<-goal_density_y$y+shots_density_y$y
goal_shot_prob_y<-(goal_density_y$y/total_shots_y)


plot(goal_shot_prob_y, 
     xlim=c(0, 100), ylim=c(0,1),
     xlab='Field Width Percentage',
     ylab='Goal Probability',
     main = 'Probability of Shot resulting in a goal vs y-coordinate of shot')
```
**Figure 4.** Success Probabilities of (Top) Assists and (Bottom) Shots as functions of (Left)  distance, (Middle) x-coordinate, and (Right) y-coordinate, and (Bottom) Shots.

## Adding Color: Looking at Differences Between Competitions and Teams

Based on this finding, it must be the case that teams consider strategies such as planning their starting points or practicing the accuracy of their shots in order to cover a desirable distance.

Regardless of the result (goal v shot, assist v key pass), how do each team's distance and starting points compare to other teams? Let's take a look at the World Cup in particular.

```{r}
# Stack assist and key pass dataset

assist_key_pass <-rbind(assists_clean, key_passes_clean)
assist_key_pass_wc <- subset(assist_key_pass, competition=="World Cup")

str(assist_key_pass_wc)
head(assist_key_pass_wc)

# Focus on top 5 teams with most assists + key passes
freq1 <- sort(table(assist_key_pass_wc$Team.Name),decreasing=T)
print(freq1)

assist_key_pass_wc_top5 <- subset(assist_key_pass_wc, Team.Name == "Belgium" | Team.Name == "England" | Team.Name == "Croatia" | Team.Name == "Brazil" | Team.Name == "France")

jpeg(file="assist_key_pass_wc_dist.jpeg")
boxplot(assist_key_pass_wc_top5$dist ~ assist_key_pass_wc_top5$Team.Name, main = "Assist or Key Pass Distance by Team", ylab = "Distance", xlab = "Team")
dev.off()

# Stack shot and goal dataset

shots_goals <-rbind(shots_clean, goals_clean)
shots_goals_wc <- subset(shots_goals, competition=="World Cup")

str(shots_goals_wc)
head(shots_goals_wc)

# Focus on top 5 teams with most goals + shots
freq2 <- sort(table(shots_goals_wc$Team.Name),decreasing=T)
print(freq2)

shots_goals_wc_top5 <- subset(shots_goals_wc, Team.Name == "Belgium" | Team.Name == "England" | Team.Name == "Croatia" | Team.Name == "Brazil" | Team.Name == "France")

jpeg(file="shots_goals_wc_dist.jpeg")
boxplot(shots_goals_wc_top5$dist ~ shots_goals_wc_top5$Team.Name, main = "Shot or Goal Distance by Team", ylab = "Distance", xlab = "Team")
dev.off()

jpeg(file="shots_goals_wc_x.jpeg")
boxplot(shots_goals_wc_top5$x.Start ~ shots_goals_wc_top5$Team.Name, main = "X.Start by Team", ylab = "X.Start", xlab = "Team")
dev.off()

jpeg(file="shots_goals_wc_y.jpeg")
boxplot(shots_goals_wc_top5$y.Start ~ shots_goals_wc_top5$Team.Name, main = "Y.Start by Team", ylab = "Y.Start", xlab = "Team")
dev.off()

```

The plots show some differences among teams. We continue by looking at these differences using an anova test.

```{r}
# install.packages("rstatix")
library(rstatix)

# Assist and Key Pass Distance
assist_key_pass_dist <- assist_key_pass_wc_top5 %>% anova_test(dist ~ Team.Name)

capture.output(assist_key_pass_dist, file = "anova.txt", append = T)

# Shots and Goals Distance
shots_goals_dist <- shots_goals_wc_top5 %>% anova_test(dist ~ Team.Name)

capture.output(shots_goals_dist, file = "anova.txt", append = T)

# Shots and Goals X.Start
shots_goals_x <- shots_goals_wc_top5 %>% anova_test(x.Start ~ Team.Name)

capture.output(shots_goals_x, file = "anova.txt", append = T)

# Shots and Goals Y.Start
shots_goals_y <- shots_goals_wc_top5 %>% anova_test(y.Start ~ Team.Name)

capture.output(shots_goals_y, file = "anova.txt", append = T)

```
Based on the anova tests, the top 5 teams (defined by event frequencies) have some significant differences in means of distances and start coordinates for the following variables:

1. Shots + Goals Distance
2. Shots + Goals Y.Start

## Conlcuding Statement:
Based on t-tests done between Shots vs Goals, and Assists vs Key Passes, there is evidence of distance and start coordinate differences between successful and unsuccessful plays. Applying this finding to the top 5 teams in the World Cup, we see further that there are teams that are significantly different in their tendencies for distance and Y.Start when it comes to attempting a goal.

## Further Work:
Further work to be included in final report:

1. Improve how we deal with bimodal distributions  
2. Improve vizualizations:  
   a. Correlate x,y-coordinate histograms with soccer pitch  
   b. 3-D success probability plots  
3. Further exploration of data by groups:  
   a. Competition  
   b. Team  

## Figure jpeg file codes:
```{r quick visuals}

#Shots
jpeg(file="shots_start.jpeg")
plot(shots_clean$x.Start,shots_clean$y.Start,main='Start Position of Shots', xlab = 'x.Start', ylab = 'y.Start')
dev.off()

jpeg(file="shots_dist.jpeg")
hist(shots_clean$dist, main='Distance of Shots', xlab = 'Shot Distance')
dev.off()

#Goals
jpeg(file="goals_start.jpeg")
plot(goals_clean$x.Start,goals_clean$y.Start,main='Start Position of Goals', xlab = 'x.Start', ylab = 'y.Start')
dev.off()

jpeg(file="goals_dist.jpeg")
hist(goals_clean$dist, main='Distance of Shots Leading to Goals', xlab = 'Shot Distance')
dev.off()

#Assists

jpeg(file="assists_start.jpeg")
plot(assists_clean$x.Start,assists_clean$y.Start,main='Start Position of Assists', xlab = 'x.Start', ylab = 'y.Start')
dev.off()

jpeg(file="assists_dist.jpeg")
hist(assists_clean$dist, main='Distance of Assists', xlab = 'Shot Distance')
dev.off()

#Key Passes

jpeg(file="keypass_start.jpeg")
plot(key_passes_clean$x.Start,key_passes_clean$y.Start,main='Start Position of Key Passes', xlab = 'x.Start', ylab = 'y.Start')
dev.off()

jpeg(file="keypass_dist.jpeg")
hist(key_passes_clean$dist, main='Distance of Key Passes', xlab = 'Shot Distance')
dev.off()
```