---
title: "CodesRUs-michael"
authors: "Daniel Diaz, Ganesh Ghimire, Rose Mesina, and Michael Womble"
date: "March 9, 2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
library(ggplot2)
library(ggsoccer)
library(ggpubr)
library(dplyr)
library(vtable)
library(purrr)
library(MASS)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "markup", message = F, echo=F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r read csv}
match_summary <- data.frame(read.csv('./datasets/match_summary.csv'))
xkabledplyhead(match_summary)
```

In order to analyze a particular variable (e.g. # goals), we need to transform the dataset from long to wide. This way, each match-team combination will have a corresponding # goals value stored in the same column.

```{r long to wide}

#Create separate datasets for Team1 and Team2. Only keeping totals as well as the relevant team's stats

team1 <- match_summary[c('team1','team1.goals', 'team1.shots','team1.shots.target','team1.passes','team1.pass.acc','team1.possession','team1.corners','team1.offsides','team1.fouls','team1.yellows','team1.reds')]

str(team1)

colnames(team1) <- c('team', 'goals', 'shots', 'shots.target', 'passes','pass.acc', 'possession', 'corners', 'offsides', 'fouls', 'yellows', 'reds')

team2 <- match_summary[c('team2','team2.goals', 'team2.shots','team2.shots.target','team2.passes','team2.pass.acc','team2.possession','team2.corners','team2.offsides','team2.fouls','team2.yellows','team2.reds')]

str(team2)

colnames(team2) <- c('team', 'goals', 'shots', 'shots.target', 'passes','pass.acc', 'possession', 'corners', 'offsides', 'fouls', 'yellows', 'reds')

teams <- rbind(team1, team2)

```

The dataset teams now has 3,882 records that can be used to analyze number of goals and the factors that influence it. We will run various linear regression models and evaluate them against one another. For this exercise, we will only use continuous variables and then check adjusted R squared to compare models.

```{r linear regressions}

teams_cont <- subset(teams, select = -c(team))
loadPkg("leaps")
best2 <- regsubsets(goals~., data = teams_cont, nbest = 2, method='seqrep')
plot(best2, scale = "adjr2", main = "Adjusted R^2")

# The best model
adjr2_model <- lm(goals ~ shots + shots.target + pass.acc + possession + corners + offsides + fouls + reds, data = teams_cont)
summary(adjr2_model)
library(car)
vif(adjr2_model)
```

The best model for goals consists of shots, shots.target, pass.acc, possession, corners, offsides, fouls, and reds. This model, however, only has an adjusted R Squared value of 0.409. This indicates that there is some information being left out of the model that could explain variation in number of goals. The only other variables available are passes and yellows. However, those variables appear to be correlated with others already in the model (see below). 

```{r correlation}
library(corrplot)
corr <- cor(teams_cont)
corrplot(corr)
```

```{r Making training and testing datasets}
# Make training and test data sets
set.seed(1)
row.number<-sample(1:nrow(teams), 0.8*nrow(teams))
teams_train<-teams[row.number,]
teams_test<-teams[-row.number,]

# Train model
model_goals<-lm(goals ~ shots + shots.target + pass.acc + possession + corners + offsides + fouls + reds, data=teams_train)
summary(model_goals)
plot(model_goals)

# Predict
pred_goals<-predict(model_goals, newdata=teams_test)
rmse <- sqrt(sum((exp(pred_goals) - teams_test$goals)^2)/length(teams_test$goals))
c(RMSE = rmse, R2=summary(model_goals)$r.squared)

plot(teams_test$goals, pred_goals)
```

```{r cross validation}
library(caret)
set.seed(2)
train_control <- trainControl(method = "repeatedcv",
                            number = 10, repeats = 3)
model <- train(goals ~., data = teams_cont,
               method = "lm",
               trControl = train_control)
pred<-predict(model, newdata = teams_test)
rmse <- sqrt(sum((exp(pred_goals) - teams_test$goals)^2)/length(teams_test$goals))
c(RMSE = rmse, R2=summary(model_goals)$r.squared)
plot(teams_test$goals, pred)
```

```{r multinomial logistic regression}
library(caret)
require(nnet)

# create separate data set from linear model
teams_logit<-teams

# Convert goals to factor type variable
teams_logit$goals<-as.factor(teams_logit$goals)

# Create training and test data sets
index<-createDataPartition(teams_logit$goals, p=0.7, list=FALSE)
train<-teams_logit[index,]
test<-teams_logit[-index,]

# Train model
train$goals<-relevel(train$goals, ref='0')
multinomial_model.fit<-multinom(goals ~ ., data=train, MaxNWts=100000)

# Test model
test$goalsPredicted<-predict(multinomial_model, newdata = test, "class")

# Results
tab<-table(test$goals, test$goalsPredicted)
tab

round((sum(diag(tab))/sum(tab))*100,2)
```

```{r simplifying match summary}
winner<-match_summary$winner
shot_ratio<-match_summary$team1.shots-match_summary$team2.shots
shot_ratio<-round(shot_ratio, digits=2)
shot_target_ratio<-match_summary$team1.shots.target-match_summary$team2.shots.target
shot_target_ratio<-round(shot_target_ratio, digits=2)
pass_ratio<-match_summary$team1.passes-match_summary$team2.passes
pass_ratio<-round(pass_ratio, digits=2)
pass_acc_ratio<-match_summary$team1.pass.acc-match_summary$team2.passes.acc
pass_acc_ratio<-round(pass_acc_ratio, digits=2)
possession_ratio<-match_summary$team1.possession-match_summary$team2.possession
possession_ratio<-round(possession_ratio, digits=2)
corners_ratio<-match_summary$team1.corners-match_summary$team2.corners
corners_ratio<-round(corners_ratio, digits=2)
offside_ratio<-match_summary$team1.offsides-match_summary$team2.offsides
offside_ratio<-round(offside_ratio, digits=2)
foul_ratio<-match_summary$team1.fouls-match_summary$team2.fouls
foul_ratio<-round(foul_ratio, digits=2)
yellow_ratio<-match_summary$team1.yellows-match_summary$team2.yellows
yellow_ratio<-round(yellow_ratio, digits=2)
red_ratio<-match_summary$team1.reds-match_summary$team2.reds
red_ratio<-round(red_ratio, digits=2)
lists<-list(winner=winner,
            shot.ratio=shot_ratio,
            shot.ratio.target=shot_target_ratio,
            pass.ratio=pass_ratio,
            pass.acc.ratio=pass_acc_ratio,
            possessio.ratio=possession_ratio,
            corners.ratio=corners_ratio,
            offside.ratio=offside_ratio,
            foul.ratio=foul_ratio,
            yellow.ratio=yellow_ratio,
            red.ratio=red_ratio)
match_diff<-as.data.frame(do.call(cbind, (lists)))
```

```{r logistical regression using match_diff}

```